---
title: "General Social Survey"
author: "Jayline Zelaya"
date: "Year 2023"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?



$H_0$ There is no relationship between gun ownership and opinion on the death penalty.


$H_A$ There is a relationship between gun ownership and opinion on the death penalty.


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->


OWNGUN is categorical variable. CAPPUN is categorical variable. Therefor we use cat~cat.rmd helper file. 


##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")

```

Among those people who favor capitol punishment they are more likely to own a gun. AMong those who oppose capitol punishment they are less likely to own a gun.


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~OWNGUN + CAPPUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```


This summary shows Among those who own a gun 75% favor capitol punishment 25% oppose it. Among those who don't own a gun
51% favor capitol punishment and 71% oppose capitol punishment.


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


Chi-sq test says p value is .0202 which is less than 0.05 so we reject the null hypothesis.


According to fisher exact test p-value is 0.01651 which is less than 0.05 therefore we conclude there is a relationship between gun ownership and capitol punishment. People who own a gun are on average 2.3x more likely to favor capitol punishment than those who do not own a gun.



#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$ There is no difference between years of education and whether or not they vote in 2012.

$H_A$ There is a difference between years of education and whether or not they vote in 2012.


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->


EDUC is numerical variable. VOTE12 is categorical variable. Therefor we use numerical~cat.rmd helper file.



##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

Those who voted in 2012 are more educated than those who didn't vote in 2012.


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```

On average those that voted in 2012 averaged 15.7. Those who did not vote in 2012 average 13.4.


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

P-value is less than 0.05 therefore we reject null hypothesis. 

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$ A person's performance on wordsum test is not affected by their level of schooling DEGREE.

$H_A$ A person's performance on wordsum test is affected by their level of schooling DEGREE.


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

WORDSUM is a numerical variable. DEGREE is a categorical variable. Therefore we use numerical~cat.rmd helper file.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
plot(WORDSUM~DEGREE,data=GSSdata)
```

This graph shows the higher the person's education level the higher their score on wordsum test.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

Those with less than a high school education averaged 6.8 wordsum score. Those with a high school diploma averaged 7.5 wordsum score. Those with a junior college degree averaged a score of 9 on the wordsum test. Those with a bachelor's degree earned an average score of 8.8 on wordsum test. Those with a graduate degree earned a average score of 9.5.


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSSdata)
summary.aov(model1)
```

P value is less than 0.05 therefore we reject the null hypothesis.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In all 3 cases P-value was less than 0.05 therefore in all 3 cases we reject the null hypothesis.
